\section{Recently Fixed Issues with Test Setup}
\label{app:alongtheway}

\subsection{Version 2}

This sub-section describe issues/lessons from writing version 2 of
the Interop report, some 8 months after the first version.

\begin{itemize}

\item Although the interop tests were running continuously over the period,
    they were not working continuously for the entire time. We had to do
        various updates of certificates but also including of the underlying
        operating system over the period and variously broke tests. So the
        first step was to get all tests working again.

\item As part of that, the selenium tests were broken by the new operating
    system, perhaps (speculating) due to a change of path name in the virtual
        environment. Re-installing the python virtual environment fixed that.

\item There was also a change in what firefox presents to selenium when we
    try to access the JSON HTTP response from our ``echstats.php'' -
        previously we got the clean JSON via an XPath
        expression for both chromium and firefox but now firefox produces
        some additional wrapping HTML that messes things up for us. We need
        to turn that off by setting the ``devtools.jsonview.enabled'' setting
        to ``False'' in the python script.

\item Updating the rust tests was an adventure - we base our test on the
    ``examples/src/bin/ech-client.rs`` script from the rustls repo, but 
        that seemed to depend on things that weren't in ``standard'' crates
        yet, so we ended up with a mixture of code from our December 2024
        test script and what's in the current rustls repository.

\item We spent a significant amount of time trying to establish for
    sure that chromium's impatience is the cause of the reduced numbers
    for those tests.

\item We did some better analysis of the domain probe data, now that we
    have more of it.

\end{itemize}

\subsection{Version 1}

The test setup used here was developed some months before this report was
written.  Over that time, clients changed, and new test clients were added, so
we had to fix a number of issues, and validate test results before producing
this report. It may be illustrative to describe some self-caused test failures
seen as we were preparing this report, and how those were resolved:

\begin{itemize}

\item Line 2 -
    \url{https://2thenp-ng.test.defo.ie:15443/echstat.php?format=json} works
        for all clients except Chromium where it never works, yet the same
        configuration on port 443 works 100\% of the time on all clients. That
        configuration has an ECHConfigList with two entries - the first an
        x25519 key and the second a p256 key - and the same ECHConfigList value
        is published for both ports. The error here however is in the
        configuration - the targetName for the port 15443 RR is ``.'' rather
        than the hostname, so Chromium here is correct and our test setup is
        wrong. We fixed a bug in
        \url{https://github.com/defo-project/ech-dev-utils/blob/main/test-cases/test_cases_gen.py}
        addressing this on 20241208.\\

    \item Line 7 and elsewhere - fixed a bug in
        \url{https://github.com/defo-project/ech-dev-utils/blob/main/scripts/selenium_test.py}
        where expected failures from tests using the ``echstat\_check()''
        handler were being logged as test failures when they should be logged
        as expected.

    \item Line 7 and elsewhere - fixed same bug in both
        \url{https://github.com/defo-project/ech-dev-utils/blob/main/scripts/smoke_ech_py.sh}
        and
        \url{https://github.com/defo-project/ech-dev-utils/blob/main/scripts/smoke_ech_rs.sh}
        where we were assuming all non-zero returns were errors, which didn't match with
        the configured expected values. (Also tweaked a few of those expected values for
        rustls and CPython too.)

    \item Line 13 - our ``cf\_check'' result handler hadn't been changed to be
        applied with the change in Cloudflare's ECH test URL from
        ``crypto.cloudflare.com'' to ``cloudflare-ech.com'' which caused
        erroneous failure reports.

    \item Line 20 and similar - our CPython test client needed a tweak to not
        fail for tests not using the ``echstat.php'' server side script - fix
        was pushed on 20241209

    \item Line 32 - somehow certbot was no longer installed on test.defo.ie so
        we got a cert expiry for hidden.hoba.ie - re-installed certbot to fix,
        which leads to some redundancy as we end up with more than one cert for
        some names (more or less all other name on that VM are below
        *.test.defo.ie)

    \item Line 35 - our CPython test client hits a certificate validation error
        for this lighttpd server - a change to the server configuration seems
        to fix this, the change being to send the full cert chain to clients
        (browsers don't need that but our other test client do).

    \item Lines 51 and 52: these needed an edit for the expected values, which were
        wrong. (That happened more than once, with the most recent fix being
        on 2024-12-13.)

\end{itemize}

Line numbers above refer to Table~\ref{tab:itests} as before.

On 2025-01-05 we also fixed some issues with temporary files that were being
left behind. That had resulted in the ``/tmp'' partition filling up on 2024-12-31
leading to some tests failing and hence skewing results between those dates.

On 2025-01-13 we updated the packages on test.defo.ie to use OpenSSL-defo rather
than OpenSSL-sftcd. We also updated all other packages on that VM, e.g. moving
from FF 133 to FF 134. (So we need to updated the s/w section above to match.)
