\section{Introduction}

Deployments of the Transport Layer Security (TLS~\cite{rfc8446}) protocol
expose the name of the server (e.g. the web site DNS name) via the Server Name
Indication (SNI) field in the first message sent (the ClientHello).  The
Encrypted Client Hello (ECH)~\cite{ietf-tls-esni} extension to TLS is a
privacy-enhancing scheme that aims to address this leak.
This report describes the current state of ECH interoperability.
The primary audience for this document are those implementing and
deploying ECH. Secondarily, there may be lessons to learn for those
designing protocols like ECH.

In May-June 2025, together with the Wikimedia Foundation, the DEfO
project carried out a small ECH trial using Wikimedia servers.\cite{echtrial}
The high-level result of that was twofold: there were no service issues
caused by ECH-enabling servers, and ECH worked for about one third of
requests to the ECH-enabled resources. As more browsers are ECH-enabled
that figure should increase, but there were also many cases of browsers
GREASEing despite that user agent string indicating that that browser
version can really do ECH.

The Open Technology Fund (OTF, \url{https://www.opentech.fund/}) have
funded the DEfO project (\url{https://defo.ie}) to develop
ECH implementations for OpenSSL, and to otherwise encourage implementation
and deployment of ECH.
As we expect the implementation and deployment environment for ECH to change
over time, this report will be updated as events warrant and is currently
versioned based on the build-time of this PDF.

The first version of this report was published in January 2025. \cite{echinterop1}
For this version we updated all software to latest versions and re-ran tests.

The latest version can be found at
\url{https://github.com/defo-project/ech-interop-report/blob/main/ech-interop-report.pdf}.
Comments, additions or corrections are welcome. Those can be sent via email to
the author, or as issues or PRs (preferably for the latex files in the
``source2'' directory) in the github repository for this report which is
\url{https://github.com/defo-project/ech-interop-report}.
