\section{Recently Fixed Issues with Test Setup}
\label{app:alongtheway}

The test setup used here was developed some months before this report was
written.  Over that time, clients changed, and new test clients were added, so
we had to fix a number of issues, and validate test results before producing
this report. It may be illustrative to describe some self-caused test failures
seen as we were preparing this report, and how those were resolved:

\begin{itemize}

\item Line 2 -
    \url{https://2thenp-ng.test.defo.ie:15443/echstat.php?format=json} works
        for all clients except Chromium where it never works, yet the same
        configuration on port 443 works 100\% of the time on all clients. That
        configuration has an ECHConfigList with two entries - the first an
        x25519 key and the second a p256 key - and the same ECHConfigList value
        is published for both ports. The error here however is in the
        configuration - the targetName for the port 15443 RR is ``.'' rather
        than the hostname, so Chromium here is correct and our test setup is
        wrong. We fixed a bug in
        \url{https://github.com/defo-project/ech-dev-utils/blob/main/test-cases/test_cases_gen.py}
        addressing this on 20241208.\\

    \item Line 7 and elsewhere - fixed a bug in
        \url{https://github.com/defo-project/ech-dev-utils/blob/main/scripts/selenium_test.py}
        where expected failures from tests using the ``echstat\_stat()''
        handler were being logged as test failures when they should be logged
        as expected.

    \item Line 7 and elsewhere - fixed same bug in both
        \url{https://github.com/defo-project/ech-dev-utils/blob/main/scripts/smoke_ech_py.sh}
        and
        \url{https://github.com/defo-project/ech-dev-utils/blob/main/scripts/smoke_ech_rs.sh}
        where we were assuming all non-zero returns were errors, which didn't match with
        the configured expected values. (Also tweaked a few of those expected values for
        rustls and CPython too.)

    \item Line 13 - our ``cf\_check'' result handler hadn't been changed to be
        applied with the change in Cloudflare's ECH test URL from
        ``crypto.cloudflare.com'' to ``cloudflare-ech.com'' which caused
        erroneous failure reports.

    \item Line 20 and similar - our CPython test client needed a tweak to not
        fail for tests not using the ``echstat.php'' server side script - fix
        was pushed on 20241209

    \item Line 32 - somehow certbot was no longer installed on test.defo.ie so
        we got a cert expiry for hidden.hoba.ie - re-instaled certbot to fix,
        which leads to some redundancy as we end up with more than one cert for
        some names (more or less all other name on that VM are below
        *.test.defo.ie)

    \item Line 35 - our CPython test client hits a certificate validation error
        for this lighttpd server - a change to the server configuration seems
        to fix this, the change being to send the full cert chain to clients
        (browsers don't need that but our other test client do).

    \item Lines 51 and 52: these needed an edit for the expected values, which were
        wrong. (That happened more than once.)

\end{itemize}

Line numbers above refer to Table~\ref{tab:itests} as before.

