\section{Conclusions}

Based on all the above, we can reach some conclusions:

\begin{itemize}

    \item
        At the ECH protocol level, interoperability, ignoring broad ECH
        ciphersuite support, is very good. The (real) issues in getting
        interoperable wide deployment of ECH revolve around complexity in HTTPS
        RR processing.

    \item
        If deploying ECH: Keep it simple, because ECH ciphersuite handling is
        not at all flexible in many clients. Only publish a ``singleton''
        ECHConfigList in one HTTPS RR and use x25519/HKDF-SHA256/AES128 for
        that one ECHConfig. Do not use aliasMode. If rotating ECH keys, do keep
        ``old'' private values usable at servers for a duration that is a
        multiple (e.g. 3x or 5x) of the duration for which the ``current'' ECH
        public value is published in the DNS. It's probably wise to also have a
        shorter TTL on HTTPS RR values, perhaps half of the key rotation
        duration or less.

    \item
        It is very non-trivial for non-browser clients to fully handle HTTPS
        RRs.  For typical non-browser clients HTTPS requires a change from the
        very simple situation of using the system resolver for A/AAAA RRs to a
        new situation where far more complex HTTPS processing may be needed. We
        expect a possibly long transition period where such clients have very
        simplistic processing of HTTPS RRs for ECH, likely just processing a
        single HTTPS RR value. Again: deployers will need to keep it simple,
        and possibly for an extended period.

    \item
        Browsers are complex beasts. It is very likely that real browser
        behaviour will be better than indicated in our test results above,
        which all involve accessing one single URL with fully ``cold'' caches.
        Nonetheless, it seems at this time, that one could not be 100\%
        confident that ECH will always be used, and always succeed, for browser
        users, even with the most conservative deployment. However, one can
        likely be confident that in such scenarios, ECH will typically be used
        and will typically succeed.

\end{itemize}

As a last word, the author is happy to use the benefit of hindsight and
speculate that had a far simpler form of ECH been developed as a first version,
with no options at all, and had that design been deployed and iterated upon, we
would be much closer to seeing ECH widely deployed today with excellent
interoperability.  The lesson there however, is for protocol designers, and not
for those implementing and deploying the current ECH specification, who are the
main target for this document.
